Temporary Trade Barriers: \\ When Will They End?

### 2016-05-13 presentation.tex
update model and uncertainty sections

\title[Temporary Trade Barriers: When Will They End?\hspace{2.15in}\insertframenumber/\inserttotalframenumber]{Temporary Trade Barriers: \\ When Will They End?}
\author[Kristy Buzard]{\texorpdfstring{Kristy Buzard\newline Syracuse University and The Wallis Institute  \newline\url{kbuzard@syr.edu}}{Kristy Buzard}}
\date{May 13, 2016}

@@

\section{Model}
\subsection{Economic and Political Structure}
\begin{frame}{Timeline}
\pause
\begin{enumerate}
	\item {\bfseries Formation}
		\begin{enumerate}[i.]
			\pause
			\item Executives set trade policy in international agreement
		\end{enumerate}
	\pause
	\item \textbf{Ratification / Maintenance}
		\begin{enumerate}[i.]
			\pause
			\item Firms lobby legislatures to break agreement
			\pause
			\item Uncertainty is resolved
			\pause
			\item Legislatures decide to break or abide by agreement
			\pause
		\end{enumerate}
	\item \textbf{Trade War} (if agreement is broken)
	\pause
		\begin{enumerate}[i.]
			\item Firms lobby legislatures to set high trade-war tariff
			\pause
			\item Uncertainty is resolved
			\pause
			\item Legislatures decide trade-war tariff
		\end{enumerate}
\pause
	\item Private actors make production, consumption decisions
\end{enumerate}
\end{frame}

@@

\begin{frame}{Political Structure}
In each country (focus on Home):
\pause
\begin{itemize}[<+->]
	\item A Unitary Executive
		\begin{itemize}
			\item Delegated authority to make trade agreement
		\end{itemize}
		%\pause
	\item A Non-unitary Legislature
		\begin{itemize}[<+->]
			\item Can withdraw delegation, break agreement, and set trade-war tariff
			\item Susceptible to influence of lobbying
			\item Decision determined by median legislator
		\end{itemize}
%\pause
	\item A Single Lobby
		\begin{itemize}
			\item Represents import-competing sector, $X$ ($Y$ in foreign)
		\end{itemize}	
\end{itemize}

@@

\subsection{The Players}
\begin{frame}
\frametitle{Executive Branch}
Trade agreement negotiated by unitary executive:
$$W_E = \mathit{CS}_X(\tau) + \mathit{CS}_Y(\tau^*) + \ga_E \pi_X(\tau) + \pi_Y(\tau^*) + \mathit{TR}(\tau)$$

\vskip.2in
\begin{itemize}
	\item $\mathit{CS_i(\cdot)}$: consumer surplus
	\item $\pi_X(\tau)$: profits of import-competing industry
	\item $\pi_Y(\tau^*)$: profits of exporting industry
	\item $\mathit{TR}(\tau)$: tariff revenue
	\item $\ga_E$: weight on profits in the import-competing industry
\end{itemize}

\begin{frame}{Non-Unitary Legislature}
  Decisions determined by preferences of Median Legislator:
\[
  W_{\mathit{ML}} = \mathit{CS}_X(\tau) + \mathit{CS}_Y(\tau^*) + \ga(e,\ta) \pi_X(\tau) + \pi_Y(\tau^*) + \mathit{TR}(\tau)
\]
\vskip-.1in
\pause
\begin{itemize}
	\item $\ga(e,\ta)$: weight on import-competing industry profits
		\begin{itemize}
			\pause
			\item $e$: lobbying effort
			\pause
			\item $\ta$: uncertain element in determination of ML's identity
		\end{itemize}
\end{itemize}

\pause
\vskip.1in
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Assumptions on $\ga(e,\ta)$}
\begin{enumerate}
  \pause
	\item $\ga(e,\ta)$ is increasing and concave in $e$ for all $\ta \in \Theta$.
  \pause
  \item $\ga(e,\ta) \geq \ga_E \geq 1 \ \forall \ta$ 
\end{enumerate}
\end{beamerboxesrounded}
\end{frame}

@@

\section{Results}
\subsection{Results: The Certain Case}
\begin{frame}{Timeline}
\begin{enumerate}
	\item \textbf{Formation}
		\begin{enumerate}[i.]
			\item Executives set trade policy in international agreement
		\end{enumerate}
	\item \textbf{Ratification / Maintenance}
		\begin{enumerate}[i.]
			\item Firms lobby legislatures to break agreement
			\item Legislatures decide to break or abide by agreement
		\end{enumerate}
	\item {\color{gray} \textbf{Trade War} (if agreement is broken)}
		\begin{enumerate}[i.]
			\item {\color{gray} Firms lobby legislatures to set high trade-war tariff}
			\item {\color{gray} Legislatures decide trade-war tariff}
		\end{enumerate}
	\item {\color{gray} Private actors make production, consumption decisions}
\end{enumerate}
\end{frame}

\begin{frame}{Highlight: Separation of Powers}
\pause
\textbf{Legislature}
\pause
\begin{itemize}
	\item Breaks agreement if median legislator prefers $\tau^{\mathit{ad}}$ to $\tau^a$
\end{itemize}

\pause
\vskip.1in
\textbf{Lobby}
\pause
\begin{itemize}[<+->]
	\item Given the $(\tau^a,\tau^{*a})$ it faces, lobby knows what $e_b$ is required to break the agreement
 	\item Lobby pays this $e_b$ if: \hskip.2in $\pi(\tau^{\mathit{tw}}) - e > \pi(\tau^a)$
\end{itemize}

\pause
\vskip.13in
\textbf{Executives}
\pause
\begin{itemize}[<+->]
	\item Set $(\tau^a,\tau^{*a})$ to make paying $e_b$ unprofitable
		\begin{itemize}
			\item[$\Rightarrow$] $e_b=0$, agreement remains in force
		\end{itemize}
	\item High tariffs, no lobbying, no trade disruptions
\end{itemize}
\end{frame}

\subsection{Full Results}
\begin{frame}{Political Uncertainty Illustration}
\pause
An Example (Bagwell $\&$ Staiger 2005)
\pause
\begin{itemize}[<+->]
	\item $D(P_i) = 1 - P_i$
	\item $Q_X(P_X) = \frac{P_X}{2}$, $Q_Y(P_Y) = P_Y$
	\item $P_X^W = \frac{4-3\tau}{7}$, $P_X = \frac{4+4\tau}{7}$
	\item $\ga(e,\ta) = 1.25 + e^{0.2} + \ta$ 
		\begin{itemize}
			\item $\ta \sim U[-0.25,0.25]$
		\end{itemize}
	\item $\ga_E = 1$
\end{itemize}

@@

\begin{frame}{Timeline}
\begin{enumerate}
	\item {\color{gray} \textbf{Formation}}
		\begin{enumerate}[i.]
			\item {\color{gray} Executives set trade policy in international agreement}
		\end{enumerate}
	\item \textbf{Ratification / Maintenance}
		\begin{enumerate}[i.]
			\item {\color{gray} Firms lobby legislatures to break agreement}
			\item Legislatures decide to break or abide by agreement
		\end{enumerate}
	\item {\color{gray} \textbf{Trade War} (if agreement is broken)}
		\begin{enumerate}[i.]
			\item {\color{gray} Firms lobby legislatures to set high trade-war tariff}
			\item {\color{gray} Legislatures decide trade-war tariff}
		\end{enumerate}
	\item {\color{gray} Private actors make production, consumption decisions}
\end{enumerate}
\end{frame}

\begin{frame}{Legislature}
  Legislature breaks trade agreement if median legislator's utility is higher under trade war than trade agreement
  \begin{itemize}
		\item Median legislator's identity is uncertain through $\ta$
	\end{itemize}

	\pause
  \vskip.2in
  Probability that Legislature breaks agreement: \\
	\pause
  \vskip.1in
    $b(e,\tau^a,\tau^{*a},\tau^{\textit{ad}},\ta)$ probability median legislator prefers $\tau^{tw}$ to $\tau^a$ for a given $\ta$
\end{frame}

\begin{frame}{Timeline}
\begin{enumerate}
	\item {\color{gray} \textbf{Formation}}
		\begin{enumerate}[i.]
			\item {\color{gray} Executives set trade policy in international agreement}
		\end{enumerate}
	\item \textbf{Ratification / Maintenance}
		\begin{enumerate}[i.]
			\item Firms lobby legislatures to break agreement
			\item {\color{gray} Legislatures decide to break or abide by agreement}
		\end{enumerate}
	\item {\color{gray} \textbf{Trade War} (if agreement is broken)}
		\begin{enumerate}[i.]
			\item {\color{gray} Firms lobby legislatures to set high trade-war tariff}
			\item {\color{gray} Legislatures decide trade-war tariff}
		\end{enumerate}
	\item {\color{gray} Private actors make production, consumption decisions}
\end{enumerate}
\end{frame}

>>>>

\title[Temporary Trade Barriers: When Will They End?\hspace{2.35in}\insertframenumber/\inserttotalframenumber]{Temporary Trade Barriers: \\ When Will They End?}
\author[Kristy Buzard]{\texorpdfstring{Kristy Buzard\newline Syracuse University and The Wallis Institute  \newline\url{kbuzard@syr.edu}}{Kristy Buzard}}
\date{May 13, 2016}

@@

\section{Model}
\subsection{Economic and Political Structure}
\begin{frame}{Timeline}
\pause

Taking trade agreement tariff and anti-dumping duties as given,
\pause
\begin{enumerate}[<+->]
	\item \textbf{Import-competing firms lobby DOC/ITC to renew AD duties}
	\item \textbf{Uncertainty is resolved}
	\item \textbf{DOC/ITC decide whether to renew duties}
	\item Private actors make production, consumption decisions
\end{enumerate}
\end{frame}

@@

\begin{frame}{Political Structure}
In Home country (foreign is passive):
\pause
\begin{itemize}[<+->]
	\item Dept. of Commerce / Int'l Trade Commission
		\begin{itemize}[<+->]
			\item Can renew AD duties
			\item Susceptible to influence of lobbying, perhaps both direct and indirect
			\item Modeled in reduced form
		\end{itemize}
%\pause
	\item A Single Lobby
		\begin{itemize}
			\item Represents import-competing sector, $X$
		\end{itemize}	
\end{itemize}

@@

\subsection{The Players}
\begin{frame}
\frametitle{``Government''}
Renewal decision determined by complex process including DOC, ITC, pressure via other political bodies. Reduced form:
\[
  W_G = \mathit{CS}_X(\tau) + \ga(e,\ta) \pi_X(\tau) + \mathit{CS}_Y(\tau^*) + \pi_Y(\tau^*) + \mathit{TR}(\tau)
\]

\begin{itemize}
	\item $\mathit{CS_i(\cdot)}$: consumer surplus
	\item $\pi_X(\tau)$: profits of import-competing industry
	\item $\pi_Y(\tau^*)$: profits of exporting industry
	\item $\mathit{TR}(\tau)$: tariff revenue
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{``Government''}
\[
  W_G = \mathit{CS}_X(\tau) + \ga(e,\ta) \pi_X(\tau) + \mathit{CS}_Y(\tau^*) + \pi_Y(\tau^*) + \mathit{TR}(\tau)
\]

\pause
\begin{itemize}
	\item $\ga(e,\ta)$: weight on import-competing industry profits
		\begin{itemize}
			\pause
			\item $e$: lobbying effort
			\pause
			\item $\ta$: uncertain element in $G$'s preferences
		\end{itemize}
\end{itemize}

\pause
\vskip.1in
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Assumption 1}
\begin{enumerate}
  \pause
	\item $\ga(e,\ta)$ is increasing and concave in $e$ for all $\ta \in \Theta$.
\end{enumerate}
\end{beamerboxesrounded}
\end{frame}

@@

\section{Uncertainty}
\subsection{}
\begin{frame}{Timeline}
\begin{enumerate}[<+->]
	\item \textbf{Import-competing firms lobby DOC/ITC to renew AD duties}
	\item {\color{gray} \textbf{Uncertainty is resolved}}
	\item \textbf{DOC/ITC decide whether to renew duties}
	\item {\color{gray} Private actors make production, consumption decisions}
\end{enumerate}

@@

\begin{frame}{Why uncertainty?}
\pause
\textbf{Legislature}
\pause
\begin{itemize}
	\item Renews AD duties if $G$ prefers $\tau^{\mathit{ad}}$ to $\tau^a$
\end{itemize}

\pause
\vskip.1in
\textbf{Lobby}
\pause
\begin{itemize}[<+->]
	\item Given $(\tau^a,\tau^{*a})$ and $\tau^{\mathit{ad}}$, lobby knows what $e$ is required to induce renewal
 	\item Lobby pays this $e$ if: \hskip.2in $\pi(\tau^{\mathit{ad}}) - e > \pi(\tau^a)$
\end{itemize}

\pause
\vskip.1in
\textbf{In Equilibrium}
\pause
\begin{itemize}[<+->]
	\item Firms only put forth effort when they know renewal will be granted
\end{itemize}

\section{Results}
\subsection{}
\begin{frame}{Timeline}
\begin{enumerate}[<+->]
	\item \textbf{Import-competing firms lobby DOC/ITC to renew AD duties}
	\item \textbf{Uncertainty is resolved}
	\item \textbf{DOC/ITC decide whether to renew duties}
	\item Private actors make production, consumption decisions
\end{enumerate}
\end{frame}

\begin{frame}{Government}
  $G$ breaks trade agreement if its utility is higher under AD duties than trade agreement tariff
  \begin{itemize}
		\item Preferences are uncertain through $\ta$
	\end{itemize}

	\pause
  \vskip.2in
  Probability that $G$ breaks agreement: \\
	\pause
  \vskip.1in
    $b(e,\tau^a,\tau^{*a},\tau^{\textit{ad}},\ta)$ probability median legislator prefers $\tau^{ad}$ to $\tau^a$ for a given $\ta$ and $e$
\end{frame}
